---
title: "Loss Leading as an Exploitative Practice"
subtitle: "*(Chen and Rey 2012)*"
author: "Alejandro Mizrahi"
date: "2024-04-09"
output: pdf_document
---

This document describes the steps to generate Figure 1.

# Analytical Solutions 

## Baseline

In the baseline, the firm L has monopoly power over A and B. This is derived in the main text:

$$\pi_{AB}^m=\frac{w_{AB}^2}{4}$$

## Case (i): $w_{AB}<w_B'$

In this scenario, as discussed in the main text, firm L exploits its monopoly power over A. Since $w_B'>w_A$, a subset of consumers who purchase B from firm S will also purchase A from L. If firm L applies zero margin, $r_A=0$, then the share of consumers is $\min\{ w_B'/2, w_A\}$. Let us consider the two cases:

- If $w_A<w_B'/2$, then demand is $w_A-r_A$. Profits are:
$$\pi_A^m:=\max_{r_A}r_A(w_A-r_A)$$

- If $w_A\geq w_B'/2$, then demand at zero margin is $w_B'/2$. The optimal margin is at least $w_A-w_B'/2$, because it does not reduce demand until this level. Profits are:
$$\pi_A^m:=r_A(w_A-[r_A-(w_A-w_B'/2)])=r_A(w_A-r_A)$$

From the FOC, $r_A^m=w_A/2$, so that $\pi_A^m=w_A^2/4$.

## Case (ii): $2w_B' \leq w_{AB}$

As derived in the main text:
$$\pi_{AB}^*=\pi_{AB}^m + \frac{(c_B-c_B')^2}{4}>\pi_{AB}^m$$

## Case (iii): $w_B'\leq w_{AB}<2w_B'$

As derived in the main text:
$$\tilde{\pi}_{AB}^*=(w_{AB}-w_B')w_B' + \frac{(c_B-c_B')^2}{4}$$

Since profits are increasing in $w_{AB}$, conceptually consider that $u_A$ increases *ceteris paribus*, we know that there is a level $w_{AB}$ such that profits match the baseline level $\pi_{AB}^m$. This occurs when $w_{AB}=w_B'+w_B$, graphically, at the point where the two lines cross. To derive this midpoint, consider that $w_{AB}=w_B'+\tau, \tau \in [0,w_B']$. Then, equating both profits:
$$\pi_{AB}^m=\tilde{\pi}_{AB}^* \Leftrightarrow \frac{w_{AB}^2}{4}=(w_{AB}-w_B')w_B'+\frac{(w_B'-w_B)^2}{4}$$
$$ \Leftrightarrow \frac{(w_{B}'+\tau)^2}{4}=\tau w_B'+\frac{(w_B'-w_B)^2}{4}$$
$$ \Leftrightarrow  \frac{\tau^2}{4}+\frac{2\tau w_B'}{4}=\tau w_B'+\frac{w_B^2}{4}-\frac{2w_B' w_B}{4}$$
$$  \Leftrightarrow \frac{\tau^2}{4}=\frac{2\tau w_B'}{4}+\frac{w_B^2}{4}-\frac{2w_B' w_B}{4}$$
$$ \Leftrightarrow \tau^2-2w_B'\tau+w_B(2w_B-w_B')=0$$

The roots of this quadratic equation are:
$$\tau=w_B' \pm (w_B'-w_B)=\begin{cases}2w_B'-w_B>w_B'\\w_B \end{cases}$$
Only the smaller (second) root is admissible in this third case, so that $\tilde{\pi}_{AB}^*=\pi_{AB}^m$ whenever $w_{AB}=w_B'+w_B$.

# Simulation

Parameters:

```{r}
cA <- 0 
cB.p <- 0 
uA.0 <- 1 
uA.1 <- 5 # st. uA.1>2*uB
uB <- 2 # st. uA.0 < uB < uA.1
min.wB <- .5 # 0<wB<uB
```

Vectors:

```{r}
cB.seq <- seq(from=cB.p,to=uB-min.wB,by=.02)
uA.seq <- seq(from=uA.0,to=uA.1,by=.02)
cB <- c(rep(uB-min.wB,length(uA.seq)),rev(cB.seq))
uA <- c(uA.seq,rep(uA.1,length(cB.seq)))

wA <- uA - cA
wB.p <- uB - cB.p
wB <- uB - cB
wAB <- wA + wB
```

Profit functions:

```{r}
piA.m <- ifelse(wAB<=wB.p,(wA^2)/4,NA)
piAB.m <- (wAB^2)/4
piAB.star <- ifelse(wAB>=2*wB.p, piAB.m + (cB-cB.p)^2/4,NA)
piAB.tilde <- ifelse(wAB>=wB.p & wAB<=2*wB.p, (wAB-wB.p)*wB.p + (cB-cB.p)^2/4,NA)
```

Raw plot:

```{r, fig.length=15,fig.width=6}
ymax <- max(piAB.star,na.rm=T)

# case 1
plot(wAB,piA.m,xlim=c(min(wAB),max(wAB)),ylim=c(0,ymax),ty="l")
# case 2
lines(wAB,piAB.star)
# case 3
lines(wAB,piAB.tilde)
# baseline
lines(wAB,piAB.m,lty=2)
# thresholds
abline(v=wB.p,lty=2,col="red")
abline(v=2*wB.p,lty=2,col="red")
```

Adjusted plot:

```{r, fig.length=15,fig.width=6}
plot(wAB,piA.m,xlim=c(min(wAB),4.5),ylim=c(0,6),ty="l",ylab="profits",
     #xaxt="n",yaxt="n"
     )
# case 2
lines(wAB,piAB.star)
# case 3
lines(wAB,piAB.tilde)
# baseline
lines(wAB,piAB.m,lty=3)
# cases
abline(v=wB.p,lty=3,col="red")
abline(v=2*wB.p,lty=3,col="red")
# legend
legend("top", legend="baseline", lty=3,cex=.7)
text(1.75, 3.5, " Case (i)",cex=.7)
text(1.75, 3, "wAB<wB'",cex=.7)
text(4.3, 3.5, " Case (ii)",cex=.7)
text(4.3, 3, "wB'<wAB<2wB'",cex=.7)
text(2.5, 3.5, " Case (iii)",cex=.7)
text(2.5, 3, "2wB<2wAB'",cex=.7)
```

```{r,results=F,echo=F}
# save plot
pdf(file="fig1.pdf",
width=8, height=6)
  plot(wAB,piA.m,xlim=c(min(wAB),4.5),ylim=c(0,6),ty="l",ylab="profits",
       #xaxt="n",yaxt="n"
       )
  # case 2
  lines(wAB,piAB.star)
  # case 3
  lines(wAB,piAB.tilde)
  # baseline
  lines(wAB,piAB.m,lty=3)
  # cases
  abline(v=wB.p,lty=3,col="red")
  abline(v=2*wB.p,lty=3,col="red")
  # legend
  legend("top", legend="baseline", lty=3,cex=1)
  text(1.75, 3.5, " Case (i)",cex=1)
  text(1.75, 3, "wAB<wB'",cex=1)
  text(4.3, 3.5, " Case (ii)",cex=1)
  text(4.3, 3, "2wB'<wAB",cex=1)
  text(2.5, 3.5, " Case (iii)",cex=1)
  text(2.5, 3, "wB'<wAB<2wB'",cex=1)
dev.off()
```